QTL_interval_plotting

### 2021-01-30 R/Analysis_v0.6_QTL_GxE_and_mash_comparison.Rmd
Finish manuscript v0.6 Intro-Discussion and four Figures

```{r}
get_post_weights <- function(m){
  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      rownames_to_column(var = "value") %>%
      mutate(value = as.integer(.data$value)) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      dplyr::select(-.data$value)

@@

  ggman_df <- pw_df %>%
    #left_join(m_meta) %>%
    separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
             extra = "merge", convert = TRUE) %>%
    left_join(log10bf_df, by = "Marker") %>%
    arrange(.data$Chr, .data$Pos) 
  return(ggman_df)
}

@@

enrich_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = qtlplot, 0.99) %>%
  mutate(subpop = "Gulf", phenotype = "flowering")
enrich_mf <- get_QTL_enrichment(m_midw_fl_hyp, qtl_df = qtlplot, 0.99) %>%
  mutate(subpop = "Midwest", phenotype = "flowering")
enrich_gamf <- get_QTL_enrichment(m_gam_fl_hyp, qtl_df = qtlplot, 0.99) %>%
  mutate(subpop = "Both", phenotype = "flowering")

@@

This random df just needs a CHR, POS_lo, and POS_hi column to work in the get_QTL_enrichment function I made. 
```{r}
sig_random <- c()

for(j in 1:100){
  qtl_size <- qtlplot %>%
    select(CHR, POS_lo, POS_hi) %>%
    unique() %>%   # don't doublecount completely overlapping QTL
    #(does doublecount partially overlapping QTL, however.)
    group_by(CHR, POS_lo) %>%
    mutate(Size = POS_hi - POS_lo) %>%
    select(CHR, POS_lo, Size)

  chr_size_mash <- pw_3f_all %>%
    group_by(Chr) %>%
    summarise(Pos_lo = min(Pos),
              Pos_hi = max(Pos))

@@

  sig_random[j] <- nrow(filter(random_df, pvalue < 0.05))
}
saveRDS(sig_random, file = "../data/Random_QTL_df_significant_enrichments_of_mash_1per_tail_SNPs_100runs.rds")
tibble(`Significant Enrichments` = sig_random) %>%
  ggplot(aes(x = `Significant Enrichments`)) +
  switchgrassGWAS::theme_oeco +
  geom_histogram(binwidth = 1) + 
  xlim(c(1, 24)) + ylab("") +
  geom_vline(xintercept = 23, linetype = 2, color = "red")
save_plot(filename = "../manuscript/plots/Figure_4_QTL_overlaps_random_QTL_enrichments_with_mash.png", plot = last_plot(), base_height = 2)
```

>>>>

```{r}
get_post_weights <- function(m){
  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      rownames_to_column(var = "value") %>%
      dplyr::mutate(value = as.integer(.data$value)) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      dplyr::select(-.data$value)

@@

  ggman_df <- pw_df %>%
    #left_join(m_meta) %>%
    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
             extra = "merge", convert = TRUE) %>%
    dplyr::left_join(log10bf_df, by = "Marker") %>%
    dplyr::arrange(.data$Chr, .data$Pos) 
  return(ggman_df)
}

@@

qtlplot <- qtlplot %>%
  arrange(desc(lod)) %>%
  mutate(pos = as.integer(pos)) %>%
  group_by(CHR, pos) %>%
  slice(1) %>%
  arrange(CHR, pos)
enrich_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = qtlplot, 0.99) %>%
  mutate(subpop = "Gulf", phenotype = "flowering")
enrich_mf <- get_QTL_enrichment(m_midw_fl_hyp, qtl_df = qtlplot, 0.99) %>%
  mutate(subpop = "Midwest", phenotype = "flowering")
enrich_gamf <- get_QTL_enrichment(m_gam_fl_hyp, qtl_df = qtlplot, 0.99) %>%
  mutate(subpop = "Both", phenotype = "flowering")

@@

This random df just needs a CHR, POS_lo, and POS_hi column to work in the get_QTL_enrichment function I made. 
```{r}
sig_random <- c()
chr_size_mash <- pw_3f_all %>%
  ungroup() %>%
  group_by(Chr) %>%
  dplyr::summarise(Pos_lo = min(Pos),
                   Pos_hi = max(Pos))

for(j in 1:1000){
  qtl_size <- qtlplot %>%
    select(CHR, POS_lo, POS_hi) %>%
    unique() %>%   # don't doublecount completely overlapping QTL
    #(does doublecount partially overlapping QTL, however.)
    group_by(CHR, POS_lo) %>%
    dplyr::mutate(Size = POS_hi - POS_lo) %>%
    dplyr::select(CHR, POS_lo, Size)

@@

  sig_random[j] <- nrow(filter(random_df, pvalue < 0.05))
}
saveRDS(sig_random, file = "../data/Random_QTL_df_significant_enrichments_of_mash_1per_tail_SNPs_1000runs.rds")
tibble(`mash 1% tail significant enrichments` = sig_random) %>%
  ggplot(aes(x = `mash 1% tail significant enrichments` )) +
  switchgrassGWAS::theme_oeco +
  geom_histogram(binwidth = 1) + 
  xlim(c(1, 24)) + ylab("") +
  geom_vline(xintercept = 16, linetype = 2, color = "red")
save_plot(filename = "../manuscript/plots/Figure_4_QTL_overlaps_random_QTL_enrichments_with_mash_1000runs.png", plot = last_plot(), base_height = 1.8, base_asp = 1.8)

tibble(sig_r = sig_random) %>%
  arrange(desc(sig_r))
3/1000
```

### 2021-01-30 R/Analysis_v0.6_QTL_interval_plotting.Rmd
Finish manuscript v0.6 Intro-Discussion and four Figures

I could combine this with some plots pulled from the mash object using `mash_plot_effects` to do a side-by-side comparison for the paper. That would go well maybe with the Figure 4 I started today, although of course I can't think of a good way to validate this statistically any more than I did in the other script. 
```{r}
library(switchgrassGWAS)
mash_plot_effects()
```

>>>>

I could combine this with some plots pulled from the mash object using `mash_plot_effects` to do a side-by-side comparison for the paper. That would go well maybe with the Figure 4 I started today, although of course I can't think of a good way to validate this statistically any more than I did in the other script. 
```{r}
library(switchgrassGWAS)
m_gulf_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_Uhyp.rds")
m_midw_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_Uhyp.rds")
m_gam_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")

get_post_weights <- function(m){
  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      rownames_to_column(var = "value") %>%
      dplyr::mutate(value = as.integer(.data$value)) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      dplyr::select(-.data$value)

  pw_df <- as_tibble(m$posterior_weights)
  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
    bind_cols(pw_df)

  ggman_df <- pw_df %>%
    #left_join(m_meta) %>%
    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
             extra = "merge", convert = TRUE) %>%
    dplyr::left_join(log10bf_df, by = "Marker") %>%
    dplyr::arrange(.data$Chr, .data$Pos) 
  return(ggman_df)
}

pw_gf <- get_post_weights(m = m_gulf_fl_hyp) %>%
  mutate(subpop = "Gulf",
         phenotype = "flowering")
pw_mf <- get_post_weights(m_midw_fl_hyp) %>%
  mutate(subpop = "Midwest",
         phenotype = "flowering")
pw_gamf <- get_post_weights(m_gam_fl_hyp) %>%
  mutate(subpop = "Both",
         phenotype = "flowering")
```

```{r}
qtldf <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv")

qtldf <- qtldf %>%
  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", convert = TRUE) %>%
  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", convert = TRUE) %>%
  mutate(Pos_Mb_lo = round(Pos_Mb_lo, digits = 2),
         Pos_Mb_hi = round(Pos_Mb_hi, digits = 2)) %>%
  select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
  arrange(CHR, Pos_Mb_lo)

qtlplot <- qtldf %>%
  mutate(phe = case_when(lodcolumn == "FL50" ~ "50% flowering",
                              lodcolumn == "dyln_fl50" ~ "Daylength",
                              lodcolumn == "dyln_change_sec" ~ "daylength change (s)",
                              lodcolumn == "CGDD_12C" ~ "GDD, GR to FL",
                              lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
  #filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
  mutate(POS_lo = Pos_Mb_lo * 1000000,
         POS_hi = Pos_Mb_hi * 1000000)

qtlplot$phe <- factor(qtlplot$phe, 
                      levels = c("50% flowering", "Rainfall, GR to FL",
                                 "GDD, GR to FL", "Daylength",
                                          "daylength change (s)",
                                          "Rainfall, 1 day",
                                                  #"Rainfall, 2d sum",
                                                  "Rainfall, 3d sum",
                                                  "Rainfall, 5d sum"#,
                                                  #"Rainfall, 7d sum"
                                                  ))
```

```{r}
pw_3f <- pw_gf %>%
  full_join(pw_mf) %>%
  full_join(pw_gamf) %>%
  arrange(Chr, Pos) %>%
  select(Chr:Pos, log10BayesFactor, subpop, value, phenotype, everything()) %>%
  filter(log10BayesFactor > 2)
pw_list <- list(Gulf = pw_gf, Midwest = pw_mf, Both = pw_gamf)
m_list <- list(Gulf = m_gulf_fl_hyp, Midwest = m_midw_fl_hyp, Both = m_gam_fl_hyp)
names(m_list)
qtlplot <- qtlplot %>%
  arrange(desc(lod)) %>%
  mutate(pos = as.integer(pos)) %>%
  group_by(CHR, pos) %>%
  slice(1) %>%
  arrange(CHR, pos)
mash_plot_manhattan_by_condition(m_gulf_fl_hyp)
mash_plot_effects(m_midw_fl_hyp, n = 5)
```

Make 4 plots for Figure 4: 
i = 4, j = 1 Both BF 7
10, 3 BF 4.37 Gulf

15, 5: 5.1 BF Midwest
16, 1 7.6 BF Both
17, 1 7.6 BF Both (same SNP as 16)
```{r}
i <- c(4, 10, 15, 16)
j <- c(1, 3, 5, 1)
f4df <- tibble(i = i, j = j)

ggdf_list <- list()

k=3
for(k in 1:nrow(f4df)){
  pw_int <- pw_3f %>%
    filter(.data$Chr %in% qtlplot$CHR[f4df$i[k]] & between(.data$Pos, 
                                                   qtlplot$POS_lo[f4df$i[k]], 
                                                   qtlplot$POS_hi[f4df$i[k]])) %>%
    arrange(desc(log10BayesFactor))

  plot_df <- 
    mash_plot_effects(m = m_list[[which(names(m_list) == pw_int$subpop[f4df$j[k]])]], 
                    i = pw_int$value[f4df$j[k]])
  ggb1 <- plot_df$effect_df %>%
    mutate(Pos_plot = plot_df$marker,
           subpop = pw_int$subpop[f4df$j[k]]) %>%
    mutate(Conditions = str_remove(Conditions, "Gulf_and_Midwest_")) %>%
  mutate(Conditions = str_remove(Conditions, "Gulf_")) %>%
  mutate(Conditions = str_remove(Conditions, "Midwest_")) %>%
  separate(Conditions, into = c("Phenotype", "Garden"), sep = -3) %>%
  mutate(Phenotype = str_remove(Phenotype, "_$"),
         Garden = str_remove(Garden, "^_")) %>%
  mutate(Garden = str_replace(Garden, "0_M", "MI"),
         Phenotype = str_replace(Phenotype, "FL5$", "FL50"))
  ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
  ggdf_list[[k]] <- ggb1
}

ggdf <- ggdf_list[[1]] %>%
  full_join(ggdf_list[[2]]) %>%
  full_join(ggdf_list[[3]]) %>%
  full_join(ggdf_list[[4]])

ggdf %>%
  ggplot() + 
  geom_point(mapping = aes(x = Garden, y = .data$mn)) +
  switchgrassGWAS::theme_oeco +
  geom_errorbar(mapping = aes(ymin = .data$mn - .data$se,
                              ymax = .data$mn + .data$se,
                              x = Garden), width = 0.3) +
  geom_hline(yintercept = 0, lty = 2) +
  facet_wrap(~Pos_plot+subpop, nrow = 1) +
  labs(x = "", y = "Effect Size (% total variation)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  coord_flip()

save_plot(filename = "../manuscript/plots/Figure_4_mash_SNP_effects.png", plot = last_plot(), base_height = 2.4, base_width = 5.5)
```
